---
title: "Extracting Multi-mappers"
author: "G.Robertson"
date: "30/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Working with bash

### Got .sam files, want to extract specific parts of the file

To extract unique mappers:
```{bash}
fgrep -w NH:i:1 seq1.5F1A_unp.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam > test.sam
```

To extract multimappers:
```{bash}
fgrep -vw NH:i:1 file > newfile
```

_________________________________________________________________________
Retain only uniquely mapping reads (reads with a single unambigous mapping location):
```{bash}
samtools view reads.bam | grep 'XT:A:U' | samtools view -bS -T referenceSequence.fa - > reads.uniqueMap.bam
```
_________________________________________________________________________


List the sam files:
```{r}
ls *.sam
```

List and Key:
0)seq1.5F1A_unp.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam
1)seq1.5F1B_p.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam
2) seq1.9F2_p.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam
3) seq2.5F1A_unp.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam
4) seq2.5F1A_unp.NG-8304_Dscam1_lib80429_4024_1_Aligned.sortedByCoord.out.sam
5) seq2.5F1A_unp.NG-8304_Dscam1_lib80429_4049_1_Aligned.sortedByCoord.out.sam
6) seq2.5F1B_p.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam
7) seq2.5F1B_p.NG-8304_Dscam1_lib80429_4024_1_Aligned.sortedByCoord.out.sam
8) seq2.5F1B_p.NG-8304_Dscam1_lib80429_4049_1_Aligned.sortedByCoord.out.sam
9) seq2.9F2_p.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam
10) seq2.9F2_p.NG-8304_Dscam1_lib80429_4024_1_Aligned.sortedByCoord.out.sam
11) seq2.9F2_p.NG-8304_Dscam1_lib80429_4049_1_Aligned.sortedByCoord.out.sam

Exclude chromosome 4:
```{bash}
fgrep -vw 4 seq1.5F1A_unp.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam > t1.txt

fgrep -vw "\*" t1.txt
```

For each .sam file:
```{bash}
fgrep -vw 4 seq1.5F1B_p.NG-8304_Dscam1_lib80429_3962_1_Aligned.sortedByCoord.out.sam > t2.txt

fgrep -vw NH:i:0 t2.txt > res1.txt
```

Each one was then merged into one file with headings indicating the key (above).

To extract only the sequence information from the above file for uploading to BLAT/BLAST:
```{bash}
# Separate line by line
xargs -n 1 <all_MM_seqs.txt  > linebyline.txt

# Extract the nt sequences
awk '/A/ && /G/ && /C/ && /T/' linebyline.txt > mmseqs.txt

# Remove duplicates:
sort -u mmseqs.txt > uniq_mms.txt
```

